Figure 3 - ChIP

### 2024-07-09 ms_figure_3_chip.Rmd
add RNA-seq analysis of (non-)bound promoter genes

```

>>>>

## Bound promoters

In `txannot$prom.number.overlapping.highconf.peaks`, we record how many TF peaks
are found in the promoter regions of a gene (defined as a 1kb window
centered on the annotated transcript start site).

Here we compare the expression levels (obtained from `rnaseq_del`) of
genes twith different numbers of TFs bound at their promoters.

```{r bound-promoters}
#| fig.width: 6
#| fig.height: 4

# calculate average expression of each gene
avg_tpm <- rowMeans(assay(rnaseq_del, "abundance"))

# add average expression to gene annotation and filter out
# genes that are not protein coding genes
pd <- txannot |>
    mutate(avg_tpm = avg_tpm[unique_einprot_id]) |>
    filter(!is.na(avg_tpm),
           gene_biotype == "protein_coding") |>
    mutate(prom.number.overlapping.highconf.peaks.categories = 
               cut(prom.number.overlapping.highconf.peaks,
                   breaks = c(-Inf, 0, 1, Inf),
                   labels = c("0", "1", ">1"),
                   right = TRUE))

# visualize
ggplot(pd, aes(prom.number.overlapping.highconf.peaks.categories, avg_tpm + 0.01)) +
    geom_violin(draw_quantiles = 0.5) +
    scale_y_log10() +
    labs(x = "Number of promoter TF peaks",
         y = "RNA level (TPM + 0.01, log-scale)") +
    theme_cowplot()
```
